---
title: "p8105_hw5_pm3260"
author: "Pradeeti Mainali"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggplot2)
library(purrr)
set.seed(101)
```

# Problem 1: Practice



# Problem 2: Explore Power

```{r sim_mu_p}

sim_q2 = function(n = 30, mu = 0, sigma = 5) {
  
  sim_data = rnorm(n, mu, sigma)
  
  t_test = broom::tidy(t.test(sim_data, mu = 0))
  
  out_df =
    tibble(
      samp_mean = mean(sim_data),
      p_value = t_test$p.value 
    )
  # t_test$p.value is the name of the variable that will be made with the previous code. 
}

output = vector("list", 5000)

for (i in 1:5000) {
  
  output[[i]] = sim_q2()
  
}

results_df = bind_rows(output)
```

`results_df` has two columns with sample mean and p-value for each of the 5000 data sets. 

```{r repeat}
results2_df = 
  expand_grid(
    mu = c(1, 2, 3, 4, 5, 6),
    iter = 1:5000
  ) |> 
  mutate(results2_df = map(mu, ~sim_q2(n = 30, mu = .x, sigma = 5))) |> 
  unnest(results2_df) 
```

created a data set with sample mean and power for mu of 1,2,3,4,5 and 6. This data set has 30,000 observations and four variables: mu, iter, samp_mean, p_value

```{r plot1}
power_res = 
  results2_df |>
  group_by(mu) |>
  summarize(power = mean(p_value < 0.05))

  ggplot(power_res, aes(x = mu, y = power)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Power of the Test vs. True Value of µ",
    x = "True Value of µ",
    y = "Power (Proportion of Null Rejections)"
  )
```

As the true value of mu increases, power also increases. This graph shows a positive association between effect size and power (larger effect sizes are more detectable leading to a higher power)

```{r plot2}
avg_mean_df = 
  results2_df |>
  group_by(mu) |>
  summarize(
    avg_mu = mean(samp_mean),
    avg_mu_rej = mean(samp_mean[p_value < 0.05])
  )

ggplot(avg_mean_df, aes(x = mu)) +
  geom_line(aes(y = avg_mu, color = "Overall Average"), size = 1.2) +
  geom_line(aes(y = avg_mu_rej, color = "Average When Null Rejected"), size = 1.2) +
  labs(
    title = "Average Estimate of Mu Hat vs. True Mu",
    x = "True Value of Mu",
    y = "Average Estimate of Mu Hat",
    color = "Estimate Type"
  ) +
  scale_color_manual(values = c("Overall Average" = "blue", "Average When Null Rejected" = "red"))
```

The sample average of mu hat across tests where the null was rejected is not approximately equal to the true mu. This is especially for smaller values of mu, where the mu hat was greater than the true mu for 1,2,3 and 4. This is likely due to selection bias. This bias diminishes as mu increases, as the effect size makes it easier to reject the null hypothesis across a broader range of sample.

# Problem 3: Homicides

```{r raw}
homicide_df = 
  read.csv("data/homicide-data.csv") |>
  janitor::clean_names()
```

The raw data contains about 52k observations and 12 variables, including uid, reported date, 

```{r citystate_md}
homicide_sum =
  homicide_df |>
  mutate(city_state = str_c(city, ", ", state),
         unsolved = disposition %in% c("Closed without arrest", "Open/No arrest")) |>
  group_by(city_state) |>
  summarize(
    total_homicides = n(),
    unsolved_homicides = sum(unsolved, na.rm = TRUE))


unsolved_prop = function(city_st) {
 
  city_data =
  homicide_sum |>
  filter(city_state == city_st)
  
  test_result = prop.test(x = pull(city_data, unsolved_homicides),
                          n = pull(city_data, total_homicides))
  tidy_result = 
    broom::tidy(test_result) |>
    select(estimate, conf.low, conf.high)
  
  return(tidy_result)
}

baltimore_results = unsolved_prop("Baltimore, MD")

```

```{r allcities}
allcities_results = 
  homicide_sum |>
  mutate(
    test_result = map2(
      unsolved_homicides,
      total_homicides,
      ~ prop.test(x = .x, n = .y)
    ),
    tidy_result = map(test_result, ~ broom::tidy(.x))
  ) |>
  unnest(tidy_result) |>
  select(city_state, estimate, conf.low, conf.high)

print(allcities_results)
```


#### Raw data description

#### Total number of homicides and unsolved homicides

#### MD proportion of homicides that are unsolved

#### Plot: Estimates and CIs for each city